Identifying Software Performance Changes Across Variants and Versions

### 2020-08-07 paper.tex
removed al wronly set apostrophes

	For each parameter combination, we record the number of measurements at each iteration, the required number of iteration for termination, as well as each iteration‘s candidate solution. To assess the accuracy of a parameter combination‘s outcome (as well as intermediate iterations), we use the $\text{F}_1$ score, a combination of the metrics precision and recall. \emph{Precision} refers to the fraction of correctly identified associations (true positives) among the associations of the retrieved (candidate) solution. \emph{Recall} is the fraction of total number of the relevant associations (i.e., those we intend to find). The $\text{F}_1$ score is the harmonic mean of precision ($P$) and recall ($R$), defined as $\text{F}_1 = 2\cdot\frac{P \cdot R}{P + R}$.	In our context, we defined a correctly change point when a commits falls in a narrow 5 commit interval from the ground truth.
	To assess efficiency, we employ the required number of iterations for termination and required measurements in relation to the $\text{F}_1$ score as proxy metrics.

	\subsection{Real-World Experiment Setup}
	In the second experiment, we evaluate our approach from a practitioner‘s perspective, where the properties and whereabouts of change points are unknown.  By means of three real-world configurable software systems, we investigate practical challenges and to determine whether our algorithm is able to detect performance-relevant commits with respect to configuration options.

@@

	\subsubsection{Operationalization}
	For the experiment, we run our approach on the performance measurements obtained in our pre-study with a few adjustments.
	First, although our study provides a broad and representative sample set of configurations, we cannot arbitrarily sample valid configurations. Instead, when acquiring new configurations (cf. Sec.~\ref{sec:acquisition}), we select configurations that (1) have not already been used by our algorithm, and (2) are closest to the requested configuration (i.e., with the minimal Hamming distance). The rationale is that this allows for rapid repetitions with different initializations. Since we employ multiple sampling strategies we are confident that our broad pre-study is representative. Second, we set the number of measurements per iteration to 300. This value is substantially smaller than in our controlled experiment setup, where we explore the scalability of our approach. Third, as an initial configuration sample set, we randomly sample 5 configurations from our pre-study set. 
	Last, we initialize our approach with a relative performance threshold $\theta_\tau$ (cf. Equation~\ref{eq:theta}) of ten percent. In the context of the reported relative variation among performance measurements on the machines used, we consider this a rather conservative threshold. 
	We repeat the experiments 10 times with different seeds to assess robustness of our approach and account for the randomness in exploring the configuration space.
	We assess practicality by reporting measurement effort, number of iterations in relation to the pre-study‘s results, which are based on a vastly greater measurement budget.

@@

	\section{Related Work}\label{sec:rw}
	\paragraph{Performance-relevant Commits} Identifying performance-changing commits is a prevalent challenge in regression testing, often applying change point detection algorithms on batches of performance observations~\cite{cityIdentifying2014,daly_industry_2020}. A way to reduce testing overhead is to employ only a fraction of performance observations or prioritize commits. Sandoval Alcocer et al. assess performance for uniformly selected commits to estimate the risk of a performance change introduced by unobserved commits~\cite{sandoval_alcocer_learning_2016,alcocer_prioritizing_2020}. Huang et al. estimate the risk of performance changes solely based on static analysis of individual commits to prioritize commits for regression testing without assessing performance~\cite{huang_performance_2014}. Mühlbauer et al. use Gaussian Processes (GP) and iterative sampling to learn performance histories with few performance observations~\cite{muhlbauer_accurate_2019}. They extract performance changing commits by applying change point detection algorithms to learned performance histories. All of the above approaches reduce testing overhead, but do not address the performance across different configurations of software systems. Our approach incorporates performance changes across different configurations by leveraging similarities in related configuration‘s performance histories.
	\paragraph{Performance of Configurable Software Systems} Associating configuration options and interactions with performance a conceptual basis for our work. There exists extensive work on modeling configuration options as features for machine learning, such as classification and regression trees~\cite{guoVariabilityawarePerformancePrediction2013,sarkarCostEfficientSamplingPerformance,nairUsingBadLearners2017}, multi-variable regression~\cite{siegmundPerformanceinfluenceModelsHighly2015}, and deep neural networks~\cite{haDeepPerf2019}. In contrast to learning performance for arbitrary configurations, our approach aims at explaining performance change, and subsequently, a change in a configuration option‘s or interaction‘s influence on performance over time. A similar scenario to ours has been studied by Jamshidi et al.~\cite{jamishidi_transfer_2017,jamshidi_transfer_gp_2017,jamshidi_learning_2018}. Between software versions, the performance influence of only few configuration options changes. The authors propose two approaches for learning a transfer function between performance models of different environments, such as versions, based on GPs~\cite{jamshidi_transfer_gp_2017} and progressive shrinking the configuration space~\cite{jamshidi_learning_2018}. We see great potential for future work in applying GPs to our problem statement as they have been successfully applied for the two orthogonal perspectives of time and configuration space.

	\section{Conclusion}
	Changes in the observed performance of a software system can be configuration-specific and difficult to pinpoint to configuration options and commits. We combine both two orthogonal perspectives, the commit history and configurability of software systems to pinpoint performance changes to individual commits and configuration options and interactions. 
	We propose an iterative sampling approach that leverages similarities in performance histories of related configurations to identify performance shifts precisely and, subsequently, pinpoint them to configuration options and interactions. 
	We have shown with a synthetic data set that our approach is sound and is able identify performance change points accurately and at scale. Experiments with three configurable software systems confirm the temporal precision of our approach and have confirmed prevalent practical challenges when studying real world-software systems, emphasizing the importance of experimental setups for evaluating algorithms. It is an avenue for further research to incorporate information, such as commit artifacts, into pinpointing performance changes as we were able to identify plausible relations of commits to configuration options to most commits with manually with little effort.

>>>>

	For each parameter combination, we record the number of measurements at each iteration, the required number of iteration for termination, as well as each iteration's candidate solution. To assess the accuracy of a parameter combination's outcome (as well as intermediate iterations), we use the $\text{F}_1$ score, a combination of the metrics precision and recall. \emph{Precision} refers to the fraction of correctly identified associations (true positives) among the associations of the retrieved (candidate) solution. \emph{Recall} is the fraction of total number of the relevant associations (i.e., those we intend to find). The $\text{F}_1$ score is the harmonic mean of precision ($P$) and recall ($R$), defined as $\text{F}_1 = 2\cdot\frac{P \cdot R}{P + R}$.	In our context, we defined a correctly change point when a commits falls in a narrow 5 commit interval from the ground truth.
	To assess efficiency, we employ the required number of iterations for termination and required measurements in relation to the $\text{F}_1$ score as proxy metrics.

	\subsection{Real-World Experiment Setup}
	In the second experiment, we evaluate our approach from a practitioner's perspective, where the properties and whereabouts of change points are unknown.  By means of three real-world configurable software systems, we investigate practical challenges and to determine whether our algorithm is able to detect performance-relevant commits with respect to configuration options.

@@

	\subsubsection{Operationalization}
	For the experiment, we run our approach on the performance measurements obtained in our pre-study with a few adjustments.
	First, although our study provides a broad and representative sample set of configurations, we cannot arbitrarily sample valid configurations. Instead, when acquiring new configurations (cf. Sec.~\ref{sec:acquisition}), we select configurations that (1) have not already been used by our algorithm, and (2) are closest to the requested configuration (i.e., with the minimal Hamming distance). The rationale is that this allows for rapid repetitions with different initializations. Since we employ multiple sampling strategies we are confident that our broad pre-study is representative. Second, we set the number of measurements per iteration to 300. This value is substantially smaller than in our controlled experiment setup, where we explore the scalability of our approach. Third, as an initial configuration sample set, we randomly sample 5 configurations from our pre-study set. 
	Last, we initialize our approach with a relative performance threshold $\theta_\tau$ (cf. Equation~\ref{eq:theta}) of ten percent. In the context of the reported relative variation among performance measurements on the machines used, we consider this a rather conservative threshold. 
	We repeat the experiments 10 times with different seeds to assess robustness of our approach and account for the randomness in exploring the configuration space.
	We assess practicality by reporting measurement effort, number of iterations in relation to the pre-study's results, which are based on a vastly greater measurement budget.

@@

	\section{Related Work}\label{sec:rw}
	\paragraph{Performance-relevant Commits} Identifying performance-changing commits is a prevalent challenge in regression testing, often applying change point detection algorithms on batches of performance observations~\cite{cityIdentifying2014,daly_industry_2020}. A way to reduce testing overhead is to employ only a fraction of performance observations or prioritize commits. Sandoval Alcocer et al. assess performance for uniformly selected commits to estimate the risk of a performance change introduced by unobserved commits~\cite{sandoval_alcocer_learning_2016,alcocer_prioritizing_2020}. Huang et al. estimate the risk of performance changes solely based on static analysis of individual commits to prioritize commits for regression testing without assessing performance~\cite{huang_performance_2014}. Mühlbauer et al. use Gaussian Processes (GP) and iterative sampling to learn performance histories with few performance observations~\cite{muhlbauer_accurate_2019}. They extract performance changing commits by applying change point detection algorithms to learned performance histories. All of the above approaches reduce testing overhead, but do not address the performance across different configurations of software systems. Our approach incorporates performance changes across different configurations by leveraging similarities in related configuration's performance histories.
	\paragraph{Performance of Configurable Software Systems} Associating configuration options and interactions with performance a conceptual basis for our work. There exists extensive work on modeling configuration options as features for machine learning, such as classification and regression trees~\cite{guoVariabilityawarePerformancePrediction2013,sarkarCostEfficientSamplingPerformance,nairUsingBadLearners2017}, multi-variable regression~\cite{siegmundPerformanceinfluenceModelsHighly2015}, and deep neural networks~\cite{haDeepPerf2019}. In contrast to learning performance for arbitrary configurations, our approach aims at explaining performance change, and subsequently, a change in a configuration option's or interaction's influence on performance over time. A similar scenario to ours has been studied by Jamshidi et al.~\cite{jamishidi_transfer_2017,jamshidi_transfer_gp_2017,jamshidi_learning_2018}. Between software versions, the performance influence of only few configuration options changes. The authors propose two approaches for learning a transfer function between performance models of different environments, such as versions, based on GPs~\cite{jamshidi_transfer_gp_2017} and progressive shrinking the configuration space~\cite{jamshidi_learning_2018}. We see great potential for future work in applying GPs to our problem statement as they have been successfully applied for the two orthogonal perspectives of time and configuration space.

	\section{Conclusion}
	Changes in the observed performance of a software system can be configuration-specific and difficult to pinpoint to configuration options and commits. We combine both two orthogonal perspectives, the commit history and configurability of software systems to pinpoint performance changes to individual commits and configuration options and interactions. 
	We propose an iterative sampling approach that leverages similarities in performance histories of related configurations to identify performance shifts precisely and, subsequently, pinpoint them to configuration options and interactions. 
	We have shown with a synthetic data set that our approach is sound and is able identify performance change points accurately and at scale. Experiments with three configurable software systems confirm the temporal precision of our approach and have confirmed prevalent practical challenges when studying real world-software systems, emphasizing the importance of experimental setups for evaluating algorithms. It is an avenue for further research to incorporate information, such as commit artifacts, into pinpointing performance changes as we were able to identify plausible relations of commits to configuration options to most commits with manually little effort.
